Co-evolution of lexical meaning \& pragmatic use

### 2017-04-04 paper/paper-v3.tex
revisions based on RHs comments, take 1 + new plots take 1

If this majority view is correct, the question arises how such a division of labor between
semantics and pragmatics could have evolved and why it would be so pervasive across natural
languages. Models of language evolution abound. There are simulation-based models studying
populations of communicating agents
\citep{Hurford1989:Biological-Evol,Steels1995:A-Self-Organizi,LenaertsJansen2005:The-Evolutionar,SteelsBelpaeme2005:Coordinating-Pe,BaronchelliPuglisi2008:Cultural-route-,steels:2011,SpikeStadler2016:Minimal-Require}
and there are mathematical models of language evolution, mostly coming from game theory
\citep{lewis:1969,Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,nowak+krakauer:1999,Huttegger2007:Evolution-and-t,Skyrms2010:Signals}. Much
of this work has focused on explaining basic properties such as compositionality and
combinatoriality
\citep[e.g.,][]{Batali1998:Computational-S,nowak+krakauer:1999,nowak+etal:2000,KirbyHurford2002:The-Emergence-o,kirby:2002,SmithKirby2003:Iterated-Learni,Gong2007:Language-Evolut,kirby+etal:2015,verhoef+etal:2014,Franke2015:Proto-Syntax},
but little attention has been paid to the interaction between conventional meaning and
pragmatic use. What is more, many mathematical models explain evolved meaning as a regularity
in the overt behavior of agents, abstracting from complex interactions between semantic
representations and pragmatic use. In contrast, we will here look at language users with a
richer cognitive make-up.

@@

\begin{example}
  Consider a simple coordination game. Agents are of two types: positive or negative. If
  agents of different type play with each other, they obtain a payoff of 0. If negative meets
  negative, each receives a payoff of 1. If positive meets positive, they get a payoff of 2. A
  population state is completely characterized by the proportion $x$ of negatives. The fitness
  of negatives in population state $x$ is $f_n(x) = x$, that of positives is $f_p(x) = 2-2x$.
  The average fitness is $\Phi(x) = x f_n(x) + (1-x) f_p(x) = 3x^2 - 4x + 2$. Without mutation,
  the replicator dynamic will then update $x$ to $RD(x) = \nicefrac{x^2}{\Phi(x)}$. The update
  function $RD(x)$ of the replicator step is plotted in Figure~\ref{fig:Updates_RMS} as the
  blue line. Rest points, for which $RD(x)=x$, are at $x=0$, $x=1$ and $x= \nicefrac{2}{3}$.
  The former are attractors as nearby points converge to them. Points near $x=\nicefrac{2}{3}$
  move towards 0 or 1. This is schematically pictured in the topmost phase portrait in
  Figure~\ref{fig:Phase_RD}.

@@

Agents play signaling games, in which the speaker wants to communicate a world state $s$ with a
message $m$ to a hearer who receives $m$ but does not know $s$
\citep[e.g.][]{lewis:1969,Skyrms2010:Signals}. A lexicon associates each message with a
(possibly fuzzy) set of states. A pragmatic type maps a lexicon onto a probabilistic speaker
rule (a probabilistic choice of message for each state) and a probabilistic hearer rule (a
probabilistic choice of state for each message). There are many ways of making these general
notions more concrete. Here is what we will assume in the remainder of this paper.

@@

Pragmatic strategies define dispositions to produce and interpret messages given a lexicon. We
distinguish between two kinds of pragmatic strategies. {\em Literal interlocutors} produce and
interpret messages literally, being guided only by their lexica. {\em Pragmatic interlocutors}
instead engage in mutual reasoning to inform their choices. Recent probabilistic models of
rational language use
\citep{frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}
capture different types of pragmatic behavior in a reasoning hierarchy. The hierarchy's bottom,
level $0$, corresponds to literal language use, as in Equations~(\ref{h:level0}) and
(\ref{s:level0}). Pragmatic language users of level $n + 1$ act (approximately) rational with
respect to level-$n$ behavior of their interlocutors, as in Equations~(\ref{h:leveln}) and
(\ref{s:leveln}).
\begin{alignat}{3}
&H_{0}(s \mid m;L) &&\propto pr(s) L_{sm} \label{h:level0}\\
&S_{0}(m \mid s;L) &&\propto \exp(\lambda \; L_{sm}) \label{s:level0}\\
&H_{n+1}(s|m;L) &&\propto pr(s) S_{n}(m|s;L) \label{h:leveln}\\
&S_{n+1}(m|s;L) &&\propto  \exp(\lambda \; H_{n}(s|m;L)) \label{s:leveln}
\end{alignat}

According to (\ref{h:level0}), a literal hearer's interpretation of a message depends on her
lexicon and her prior over states, $pr \in \Delta(S)$, which is here assumed flat for
simplicity. Literal interpreters thereby choose an arbitrary true interpretation for each
message. Pragmatic hearers, defined in (\ref{h:leveln}), instead use Bayes' rule to weigh
interpretations based on a conjecture about speaker behavior. Speaker behavior is regulated by
a soft-max parameter $\lambda \geq 0$ \citep{luce:1959,sutton+barto:1998}. As
$\lambda$ increases, choices approximate strict maximization of expected utilities. Expected
utility of a message $\messg$ in state $\state$ for a level $n+1$ speaker is here defined as
$H_{n}(s|m;L)$, the probability that the hearer will assign to or choose the correct
meaning. For literal speakers, utility only tracks truthfulness. Literal speakers choose any
true message with equal probability but may send false messages as well with a probability
dependent on $\lambda$.

@@

If $\lambda = 1$ literal speaker behavior is as follows:

@@

If $\lambda = 20$, literal speakers will choose true utterances with less slack:

@@

Pragmatic hearers of level 1 have the following probabilistic interpretation behavior for
$\lambda = 1$:
For $\lambda = 20$, there will less slack again:

@@

With $\lambda = 1$, pragmatic speakers of level 1 will have:

@@

For high $\lambda = 20$, pragmatic speakers of level 1 give us:

@@

In the following we consider a specific application of the model from Section~\ref{sec:model}
which allows to address the question if or when scalar inferences might (not) lexicalize. We
consider what is perhaps one of the simplest non-trivial setups that speak to this matter and
reflect on its limitations in Section~\ref{sec:discussion}. The setup is introduced in
Section~\ref{sec:setup:-stat-conc}. Section~\ref{sec:simulation-results} describes simulations
and their results.

@@

A lexicon $L$ is a mapping $\Messgs \rightarrow R$ from messages to representations. With three
messages there are $6^3 = 216$ possible lexica. Some assign the same representations to more
than one message and others lexicalize the same representations but associate them with
different messages. Out of these possible lexica, three kinds are of particular
relevance. First, lexica that assign the same lexical representations to more than one
message. Such lexica lack in expressivity but may be favored by particular learning biases
nonetheless (see below). Second, lexica that conventionalize upper-bounds to realize a
one-to-one mapping of messages to states. Finally, lexically that do not lexicalize an upper
bound but allow it to be conveyed pragmatically due to the presence of a stronger lexical
item. There are six lexica of the second kind and six of the third.  The following three lexica
exemplify each kind:
\begin{align*}
  \begin{blockarray}{lccc}
    & & \underline{\Lall} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
     \mygray{\snone}  & 0 & 0 & 0\\
     \mygray{\ssome}  & 0 & 0 & 0\\
    \mygray{\sall}   & 1 & 1 & 1\\
    \end{block}
% \\ & & P(\Lall)\rlap{ $\propto 9$} & \\
  \end{blockarray} & &
 \begin{blockarray}{lccc}
    & & \underline{\Lbound} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 0 & 1\\
    \end{block}
%  \\& & P(\Lbound)\rlap{ $\propto 96$} & \\
  \end{blockarray} & &
  \begin{blockarray}{lccc}
    & & \underline{\Llack} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 1 & 1\\
    \end{block}
%   \\& & P(\Llack)\rlap{ $\propto 48$} & \\
  \end{blockarray}
\end{align*}
Lexicon $\Lall$ is clearly bad for communication: all message and interpretation choices will
be equally likely for all types; no information about the observed world state will be conveyed
by its users. In contrast, users of $\Lbound$ can communicate world states perfectly, no matter
whether they are literal or pragmatic users. Users of $\Llack$ can also communicate information
about the actual world state but need pragmatic language use to approximate a one-to-one
mapping between message use and states (see Section~\ref{sec:languages+use}).

@@

With respect to (i), \citet{kirby+etal:2015} propose a comparable model of the interaction
between a lexicon's expressivity and its learnability. A main difference is that here we
considered a pressure for mutual understanding within a generation that may indirectly select
for expressive types, i.e., those that can convey states unequivocally, whereas Kirby et
al. only consider the bearing that the latter ability has on the production of learnable
data. We see two main reasons for considering communicative success rather than just
expressivity, and to only considering it in learning. First, learnability alone can lead to the
acquisition of functionally defective languages, as showcased by $\Lall$
(cf. \citealt{kirby+etal:2008,silvey+etal:2014}, and \citealt{fay+etal:2013} for review of
laboratory results). Second and more importantly, types may be equally expressive but their
performance as means of information transfer crucially depends not only on themselves but on
the population they find themselves in (cf. the competition of target types in Figure
\ref{fig:rmd} and their lack of competition in Figure \ref{fig:only-M}). That is, we contend
that having an expressive lexicon that generates learnable data does not in itself capture a
type's arguably central communicative function of transferring information to peers. Acquiring
a maximally expressive type is of little use when it fails to be understood in its population.

@@

What is more, it is not that we just assumed a prior disadvantage of lexicalized upper
bounds. We tried to motivate and formalize a general assumption about lexical representations'
complexity with a concrete, albeit provisional proposal. The specification of a learning bias
in terms of a ``grammar of representations'' can and should be seen critically, however. Much
depends on the primitives of such a grammar. For instance, the lexical representation ``none or
all'' is the most complex in Table~\ref{tab:concepts}. But consider adding a new primitive
relation between sets $A \smile B$ which is true if and only if
$\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$. The lexical representation ``none or
all'' would then be one of the simplest. Clearly, further research, empirical and conceptual,
into the role of representational complexity, processing costs and learning biases is
needed. The model here makes a clear and important contribution nonetheless: it demonstrates
how simplicity of representations can interact with use and evolutionary selection and shows
that for simple representations to emerge it may require pragmatic strategies to compensate
their potential expressive deficiencies. Hence model of co-evolving semantics and pragmatics is
needed. Future work should also include the possibility that representational simplicity may
itself be a notion that is subject to evolutionary pressure
\citep[cf.][]{ThompsonKirby2016:Culture-Shapes-}.

Finally, our case study should not be interpreted as a proposal for a definite explanation of
how scalar implicatures evolved. Other factors should be considered eventually even if they
will lead to much more complex modeling. One such factor is the observation that
non-lexicalized upper bounds allow a broader range of applicability, e.g., when the speaker is
not certain as to whether \emph{all} is true. This may suggest an alternative and purely
functionalist argument for why upper-bounded meanings do not conventionalize: should contextual
cues provide enough information to the hearer to identify whether a bound is intended to be
conveyed pragmatically, then this is preferred over expressing it overtly through longer
expressions, e.g., by saying {\em some but not all} explicitly. Importantly, although
morphosyntactic disambiguation may be dispreferred due to its relative length and complexity
\citep{piantadosi+etal:2012b}, it allows speakers to enforce an upper-bound and override
contextual cues that might otherwise mislead the hearer. In a nutshell, this explanation posits
that scalar implicatures fail to lexicalize because, all else being equal, speakers prefer to
communicate as economically as possible and pragmatic reasoning enables them to do so. What
this alternative argument does not explain is why functional pressure does not lead to the
emergence of different, equally costly lexical items to express different knowledge states of
the speaker (cf. \citealt[252-267]{horn:1984},
\citealt{horn:1972,traugott:2004,vdAuwera:2010}). For instance, two expressions for each weak
scalar expression; one with and one lacking an upper-bound.  Looking at pressure from
learnability might come in again. The present work made a first start and gave a framework for
exploring exactly these issues systematically.

@@

% \appendix

>>>>

If this majority view is correct, the question arises how such a division of labor between
semantics and pragmatics could have evolved, why it would be so pervasive across natural
languages and why it is that some expressions systematically draw from it while others semantically conventionalize. Models of language evolution abound. There are simulation-based models studying
populations of communicating agents
\citep{Hurford1989:Biological-Evol,Steels1995:A-Self-Organizi,LenaertsJansen2005:The-Evolutionar,SteelsBelpaeme2005:Coordinating-Pe,BaronchelliPuglisi2008:Cultural-route-,steels:2011,SpikeStadler2016:Minimal-Require}
and there are mathematical models of language evolution, mostly coming from game theory
\citep{lewis:1969,Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,nowak+krakauer:1999,Huttegger2007:Evolution-and-t,Skyrms2010:Signals}. Much
of this work has focused on explaining basic properties such as compositionality and
combinatoriality
\citep[e.g.,][]{Batali1998:Computational-S,nowak+krakauer:1999,nowak+etal:2000,KirbyHurford2002:The-Emergence-o,kirby:2002,SmithKirby2003:Iterated-Learni,Gong2007:Language-Evolut,kirby+etal:2015,verhoef+etal:2014,Franke2015:Proto-Syntax},
but little attention has been paid to the interaction between conventional meaning and
pragmatic use. What is more, many mathematical models explain evolved meaning as a regularity
in the overt behavior of agents, abstracting from complex interactions between semantic
representations and pragmatic use. In contrast, we will here look at language users with a
richer cognitive make-up.

@@

\begin{example}
  Consider a simple coordination game. Agents are of two types: positive or negative. If
  agents of different type play with each other, they obtain a payoff of 0. If negative meets
  negative, each receives a payoff of 1. If positive meets positive, they get a payoff of 2. A
  population state is completely characterized by the proportion $x$ of negatives. The fitness
  of negatives in population state $x$ is $f_n(x) = x$, that of positives is $f_p(x) = 2-2x$.
  The average fitness is $\Phi(x) = x f_n(x) + (1-x) f_p(x) = 3x^2 - 4x + 2$. Without mutation,
  the replicator dynamic will then update $x$ to $RD(x) = \nicefrac{f(x)x}{\Phi(x)} = \nicefrac{x^2}{\Phi(x)}$. The update
  function $RD(x)$ of the replicator step is plotted in Figure~\ref{fig:Updates_RMS} as the
  blue line. Rest points, for which $RD(x)=x$, are at $x=0$, $x=1$ and $x= \nicefrac{2}{3}$.
  The former are attractors as nearby points converge to them. Points near $x=\nicefrac{2}{3}$
  move towards 0 or 1. This is schematically pictured in the topmost phase portrait in
  Figure~\ref{fig:Phase_RD}.

@@

Agents play signaling games, in which the speaker wants to communicate a world state $s$ with a
message $m$ to a hearer who receives $m$ but does not know $s$
\citep[e.g.][]{lewis:1969,Skyrms2010:Signals}. A lexicon associates each message with a
(possibly fuzzy) set of states. A pragmatic type specifies a probabilistic speaker
rule (a probabilistic choice of message for each state) and a probabilistic hearer rule (a
probabilistic choice of state for each message) given a lexicon. There are many ways of making these general
notions more concrete. Here is what we will assume in the remainder of this paper.

@@

Pragmatic strategies define dispositions to produce and interpret messages given a lexicon. We
distinguish between two kinds of pragmatic strategies. {\em Literal interlocutors} produce and
interpret messages literally, being guided only by their lexica. {\em Pragmatic interlocutors}
instead engage in mutual reasoning to inform their choices. Recent probabilistic models of
rational language use
\citep{frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}
capture different types of pragmatic behavior in a reasoning hierarchy. The hierarchy's bottom,
level $0$, corresponds to literal language use, as in Equations~(\ref{h:level0}) and
(\ref{s:level0}). Pragmatic language users of level $n + 1$ act (approximately) rational with
respect to level-$n$ behavior of their interlocutors, as in Equations~(\ref{h:leveln}) and
(\ref{s:leveln}).
\begin{alignat}{3}
&H_{0}(s \mid m;L) &&\propto pr(s) L_{[s,m]} \label{h:level0}\\
&S_{0}(m \mid s;L) &&\propto \exp(\lambda \; L_{[s,m]}) \label{s:level0}\\
&H_{n+1}(s|m;L) &&\propto pr(s) S_{n}(m|s;L) \label{h:leveln}\\
&S_{n+1}(m|s;L) &&\propto  \exp(\lambda \; H_{n}(s|m;L)) \label{s:leveln}
\end{alignat}

According to (\ref{h:level0}), a literal hearer's interpretation of a message depends on her
lexicon and her prior over states, $pr \in \Delta(S)$, which is here assumed flat for
simplicity. Literal interpreters thereby choose an arbitrary true interpretation for each
message according to their lexicon. Pragmatic hearers, defined in (\ref{h:leveln}), instead use Bayes' rule to weigh
interpretations based on a conjecture about speaker behavior. Speaker behavior is regulated by
a soft-max parameter $\lambda \geq 0$ \citep{luce:1959,sutton+barto:1998}. As
$\lambda$ increases, choices approximate strict maximization of expected utilities. Expected
utility of a message $\messg$ in state $\state$ for a level $n+1$ speaker is here defined as
$H_{n}(s|m;L)$, the probability that the hearer will assign to or choose the correct
meaning. For literal speakers, utility only tracks truthfulness. Literal speakers choose any
true message with equal probability but may send false messages as well with a probability
dependent on $\lambda$.

@@

By contrast, pragmatic hearers of level 1 have the following interpretative behavior for $\lambda = 1$:
This is the outcome of reasoning about their level-$0$ speaker counterparts with $\lambda = 1$:

@@

With low $\lambda$ speakers choose true messages with more slack. Reasoning over this behavior therefore also results in a weaker association of messages with only true states in receivers, but also in a slightly stronger association of $\msome$ with $\ssome$ over $\sall$ for $\Llack$ users because $S_0(\msome | \ssome;\Llack) > S_0(\msome | \sall;\Llack)$. For $\lambda = 20$, there will be less slack in literal speaker behavior:

@@

And accordingly less slack in pragmatic interpretation:

@@

Lastly, turning to types that have no bearing on hearer choices, with $\lambda = 1$ pragmatic speakers of level 1 have:

@@

For $\lambda = 20$, pragmatic speaker behavior of level 1 is as follows:

@@

In the following we consider a specific application of the model from Section~\ref{sec:model}
which allows us to address the question if or when scalar inferences might (not) lexicalize. We
consider what is perhaps one of the simplest non-trivial setups that speak to this matter and
reflect on its limitations in Section~\ref{sec:discussion}. The setup is introduced in
Section~\ref{sec:setup:-stat-conc}. Section~\ref{sec:simulation-results} describes simulations
and their results.

@@

A lexicon $L$ is a mapping $\Messgs \rightarrow R$ from messages to representations. With three
messages there are $6^3 = 216$ possible lexica. Some assign the same representations to more
than one message and others lexicalize the same representations but associate them with
different messages. Out of these possible lexica, three kinds are of particular
relevance. First, lexica that assign the same lexical representations to more than one
message. Such lexica lack in expressivity but may be favored by particular learning biases
nonetheless (see below). Second, lexica that conventionalize upper-bounds to realize a
one-to-one mapping of messages to states. Finally, lexica that do not lexicalize an upper
bound but allow it to be conveyed pragmatically due to the presence of a stronger lexical
item. There are six lexica of the second kind and six of the third.  The following three lexica
exemplify each kind:
\begin{align*}
  \begin{blockarray}{lccc}
    & & \underline{\Lall} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
     \mygray{\snone}  & 0 & 0 & 0\\
     \mygray{\ssome}  & 0 & 0 & 0\\
    \mygray{\sall}   & 1 & 1 & 1\\
    \end{block}
% \\ & & P(\Lall)\rlap{ $\propto 9$} & \\
  \end{blockarray} & &
 \begin{blockarray}{lccc}
    & & \underline{\Lbound} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 0 & 1\\
    \end{block}
%  \\& & P(\Lbound)\rlap{ $\propto 96$} & \\
  \end{blockarray} & &
  \begin{blockarray}{lccc}
    & & \underline{\Llack} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 1 & 1\\
    \end{block}
%   \\& & P(\Llack)\rlap{ $\propto 48$} & \\
  \end{blockarray}
\end{align*}
Lexicon $\Lall$ is clearly bad for communication: all message and interpretation choices will
be equally likely for all types; no information about the observed world state will be conveyed
by its users. In contrast, users of $\Lbound$ can communicate world states perfectly, no matter
whether they are literal or pragmatic users. Users of $\Llack$ can also communicate information
about the actual world state but need pragmatic language use to approximate a one-to-one
mapping between message use and states (see Section~\ref{sec:languages+use}).

@@

With respect to (i), \citet{kirby+etal:2015} propose a comparable model of the interaction
between a lexicon's expressivity and its learnability. A main difference is that here we
considered a pressure for mutual understanding within a generation that may indirectly select
for expressive types, i.e., those that can convey states unequivocally, whereas Kirby et
al. only consider the bearing that the latter ability has on the production of learnable
data. We see two main reasons for considering communicative success rather than just
expressivity, and to only considering it in learning. First, learnability alone can lead to the
acquisition of functionally defective languages, as showcased by $\Lall$
(cf. \citealt{kirby+etal:2008,silvey+etal:2014}, and \citealt{fay+etal:2013} for review of
laboratory results). Second and more importantly, types may be equally expressive but their
performance as means of information transfer crucially depends not only on themselves but on
the population they find themselves in (cf. the competition of target types in Figure
\ref{fig:rmd} and their lack of competition in Figure \ref{fig:only-M}). That is, we contend
that having an expressive lexicon that generates learnable data does not in itself capture a
type's arguably central communicative function of transferring information to peers. Taking communication
into consideration in turn allows the model to be responsive to the task for which language is learned.

@@

What is more, it is not that we just assumed a prior disadvantage of lexicalized upper
bounds. We tried to motivate and formalize a general assumption about lexical representations'
complexity with a concrete, albeit provisional proposal. The specification of a learning bias
in terms of a ``grammar of representations'' can and should be seen critically, however. Much
depends on the primitives of such a grammar. For instance, the lexical representation ``none or
all'' is the most complex in Table~\ref{tab:concepts}. But consider adding a new primitive
relation between sets $A \smile B$ which is true if and only if
$\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$. The lexical representation ``none or
all'' would then be one of the simplest. Clearly, further research, empirical and conceptual,
into the role of representational complexity, processing costs and learning biases is
needed. The model here makes a clear and important contribution nonetheless: it demonstrates
how simplicity of representations can interact with use and evolutionary selection and shows
that for simple representations to emerge it may require pragmatic strategies to compensate
their potential expressive deficiencies. Hence a model of co-evolving semantics and pragmatics is
needed. Future work should also include the possibility that representational simplicity may
itself be a notion that is subject to evolutionary pressure
\citep[cf.][]{ThompsonKirby2016:Culture-Shapes-}, as well for the evolution of elements that define
the agents' cognitive make-up; $\lambda$ and $l$.

Finally, our case study should not be interpreted as a proposal for a definite explanation of
how scalar implicatures evolved. Other factors should be considered eventually even if they
will lead to much more complex modeling. One such factor is the observation that
non-lexicalized upper bounds allow a broader range of applicability, e.g., when the speaker is
not certain as to whether \emph{all} is true. This may suggest an alternative and purely
functionalist argument for why upper-bounded meanings do not conventionalize: should contextual
cues provide enough information to the hearer to identify whether a bound is intended to be
conveyed pragmatically, then this is preferred over expressing it overtly through longer
expressions, e.g., by saying {\em some but not all} explicitly. Importantly, although
morphosyntactic disambiguation may be dispreferred due to its relative length and complexity
\citep{piantadosi+etal:2012b}, it allows speakers to enforce an upper-bound and override
contextual cues that might otherwise mislead the hearer. In a nutshell, this explanation posits
that scalar implicatures fail to lexicalize because, all else being equal, speakers prefer to
communicate as economically as possible and pragmatic reasoning enables them to do so. What
this alternative argument does not explain is why functional pressure does not lead to the
emergence of different, equally costly lexical items to express different knowledge states of
the speaker (cf. \citealt[252-267]{horn:1984},
\citealt{horn:1972,traugott:2004,vdAuwera:2010}). For instance, two expressions for each weak
scalar expression; one with and one lacking an upper-bound.  Looking at pressure from
learnability might come in again. Beyond scalar implicatures, the model can generate predictions about
likely lexicalization trajectories of pragmatic inferences, or a lack thereof. In this realm
an interesting issue is whether proposed principles, such as the semantic conventionalization of once highly context-dependent inferences if they become regular enough \citep{levinson:2000,traugott:2004}, can be given
a formal rationale and inform the postulated directionalities of change. The present work made a first start and gave a framework for
exploring exactly these issues systematically.

@@

 \appendix
\newpage

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/alt-fig1-onlyr}
\caption{Representative population after $50$ generations under only a
  pressure for communicative success. The right-most plots zoom in on only the proportion of
  target types.}
\label{fig:only-R}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth,height=8cm,keepaspectratio]{./plots/alt-fig2-onlym-pr}
\caption{Representative population after $50$ generations under only a
  pressure for learnability ($\lambda = 20, k = 5$). The types' prior is shown in the
  right-most plot. The top-most groupings correspond to types, literal and pragmatic, with
  lexica of the kind of $\Llack$, $\Lbound$, and $\Lall$.}
\label{fig:only-M}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth,height=8cm,keepaspectratio]{./plots/alt-fig3-r+m}
\caption{Representative population after $50$ generations under both pressures ($k = 5$).}
\label{fig:rmd}
\end{figure}
